---
title: "Instrument Perception - English Pilot"
author: "Grace Brown"
date: "2026-01-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)

knitr::opts_knit$set(root.dir = "/Users/gracebrown/instruments-english")
setwd("/Users/gracebrown/instruments-english")

# color-blind-friendly palette
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
theme_set(theme_bw())
```

## Load Data
```{r data}
df <- read_csv("./data/pilot_combined_expanded.csv",
               show_col_types = FALSE) 
stim_key <- read_csv("./data/pilot-stim-key-english.csv",
                     show_col_types = FALSE)
```
## Sentence Guide
```{r key, echo=FALSE}
knitr::kable(stim_key)
```

## Visualization
### Natural
```{r plot natural, echo=FALSE}
plot.natural <- df %>% 
  filter(!is.na(subj)) %>%
  ggplot(mapping = aes(x = id, 
                       y = natural)) +
  stat_summary(fun = "mean",
               geom = "bar") + 
  stat_summary(fun.data = "mean_cl_boot",
               geom = "linerange") + 
  coord_cartesian(ylim = c(0,7)) +
  labs(x = "Sentence ID",
       y = "Mean Rating 'Sentence is Natural'",
       title = "Mean Naturalness Rating by Sentence and Subject") + 
  facet_grid(~subj)

print(plot.natural)
```

Overall, this is patterning as expected, although I should definitely look into item-specific differences for instrument-subject items.

### Mobile
```{r plot mobile, echo=FALSE}
plot.mobile <- df %>% 
  filter(!is.na(subj)) %>%
  ggplot(mapping = aes(x = id, 
                       y = mobile)) +
  stat_summary(fun = "mean",
               geom = "bar") + 
  stat_summary(fun.data = "mean_cl_boot",
               geom = "linerange") + 
  coord_cartesian(ylim = c(0,7)) +
  labs(x = "Sentence ID",
       y = "Mean Rating 'Was Mobile'",
       title = "Mean Mobility Rating by Sentence and Subject") + 
  facet_grid(~subj)

print(plot.mobile)
```

The mobility noise is definitely item-specific; I may need to revise these items. 

### Volition
```{r plot volition, echo=FALSE}
plot.volition <- df %>% 
  filter(!is.na(subj)) %>%
  ggplot(mapping = aes(x = id, 
                       y = volition)) +
  stat_summary(fun = "mean",
               geom = "bar") + 
  stat_summary(fun.data = "mean_cl_boot",
               geom = "linerange") + 
  coord_cartesian(ylim = c(0,7)) +
  labs(x = "Sentence ID",
       y = "Mean Rating 'Acted of Own Volition'",
       title = "Mean Volitionality Rating by Sentence and Subject") + 
  facet_grid(~subj)

print(plot.volition)
```

Patterning as expected.

### Potent
```{r plot potent, echo=FALSE}
plot.potent <- df %>% 
  filter(!is.na(subj)) %>%
  ggplot(mapping = aes(x = id, 
                       y = potent)) +
  stat_summary(fun = "mean",
               geom = "bar") + 
  stat_summary(fun.data = "mean_cl_boot",
               geom = "linerange") + 
  coord_cartesian(ylim = c(0,7)) +
  labs(x = "Sentence ID",
       y = "Mean Rating 'Generated Own Engergy to Do X'",
       title = "Mean Potency Rating by Sentence and Subject") + 
  facet_grid(~subj)

print(plot.potent)
```

Note that the later sentence IDs are the "machines" or animate non-humans, so this patterns as expected as well.

### Sentient
```{r plot sentient, echo=FALSE}
plot.sentient <- df %>% 
  filter(!is.na(subj)) %>%
  ggplot(mapping = aes(x = id, 
                       y = sentient)) +
  stat_summary(fun = "mean",
               geom = "bar") + 
  stat_summary(fun.data = "mean_cl_boot",
               geom = "linerange") + 
  coord_cartesian(ylim = c(0,7)) +
  labs(x = "Sentence ID",
       y = "Mean Rating 'Is Sentient'",
       title = "Mean Sentience Rating by Sentence and Subject") + 
  facet_grid(~subj)

print(plot.sentient)
```

Patterning as expected! Solid baseline. 

### Instigation
```{r plot instigation, echo=FALSE}
plot.instigation <- df %>% 
  filter(!is.na(subj)) %>%
  ggplot(mapping = aes(x = id, 
                       y = instigation)) +
  stat_summary(fun = "mean",
               geom = "bar") + 
  stat_summary(fun.data = "mean_cl_boot",
               geom = "linerange") + 
  coord_cartesian(ylim = c(0,7)) +
  labs(x = "Sentence ID",
       y = "Mean Rating 'Caused X'",
       title = "Mean Instigation Rating by Sentence and Subject") + 
  facet_grid(~subj)

print(plot.instigation)
```

Patterning as expected! Nice sanity check.

### Qualitative Persistence 
```{r plot qualpersist, echo=FALSE}
plot.qualpersist <- df %>% 
  filter(!is.na(subj)) %>%
  ggplot(mapping = aes(x = id, 
                       y = qualpersist)) +
  stat_summary(fun = "mean",
               geom = "bar") + 
  stat_summary(fun.data = "mean_cl_boot",
               geom = "linerange") + 
  coord_cartesian(ylim = c(0,7)) +
  labs(x = "Sentence ID",
       y = "Mean Rating 'Did Not Change During Event'",
       title = "Mean Qualitative Persistance Rating by Sentence and Subject") + 
  facet_grid(~subj)

print(plot.qualpersist)
```

Much more variation than expected here. Specifically, we would expect most (if not everything) to have a relatively high degree of qualitative persistence. Likely, the wording of the question is being interpreted variably.